Theorems and Definitions

### 2016-04-05 math_modeling/hw/hw6/hw6.tex
working on modeling
working on number theory

@@

@@

\begin{minipage}{0.5\textwidth}
\[
\begin{array}{c | c c c}
& \text{L} & \text{C} & \text{N}\\
\hline
\text{L} & .6 & .2 & .15\\
\text{C} & .2 & .8 & .1\\
\text{N} & .2 & 0 & .75
\end{array}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {L};
   \node[state] (q_1) [right=of q_0] {C};
   \node[state] (q_2) [right=of q_1] {N};
   % \node[state] (q_2) [below right=of q_0] {$q_2$};
   % \node[state,accepting](q_3) [below right=of q_1] {$q_3$};
    \path[->]
    (q_0)   edge [bend left] node {0.2} (q_1)
            edge [bend left=70] node {0.2} (q_2)
            edge [loop left] node {0.6} ()
    (q_1)   edge [bend left] node {0.2} (q_0)
            edge [loop above] node {0.8} ()
    (q_2)   edge [bend left=70] node {.15} (q_0)
            edge [bend left] node {.1} (q_1)
            edge [loop right] node {.75} ();
\end{tikzpicture}
\end{center}
\end{minipage}
\end{enumerate}

@@

\item Using probabilistic considerations only, show that the square of a stochastic matrix is also a
stochastic matrix.
\begin{quote}

\end{quote}
\end{enumerate}

>>>>

%% Quiz 4 on page 22/23

@@

\usepackage{multicol}

@@

\begin{minipage}{0.5\textwidth}
\[
\begin{array}{c | c c c}
& \text{L} & \text{C} & \text{N}\\
\hline
\text{L} & .6 & .2 & .2\\
\text{C} & .2 & .8 & 0\\
\text{N} & .15 & .1 & .75
\end{array}
\]
\end{minipage}
\begin{minipage}{0.5\textwidth}
\begin{center}
\begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
   \node[state] (q_0) {L};
   \node[state] (q_1) [right=of q_0] {C};
   \node[state] (q_2) [right=of q_1] {N};
   % \node[state] (q_2) [below right=of q_0] {$q_2$};
   % \node[state,accepting](q_3) [below right=of q_1] {$q_3$};
    \path[->]
    (q_0)   edge [bend left] node {0.2} (q_1)
            edge [bend left=70] node {0.2} (q_2)
            edge [loop left] node {0.6} ()
    (q_1)   edge [bend left] node {0.2} (q_0)
            edge [loop above] node {0.8} ()
    (q_2)   edge [bend left=70] node {.15} (q_0)
            edge [bend left] node {.1} (q_1)
            edge [loop right] node {.75} ();
\end{tikzpicture}
\end{center}
\end{minipage}
\end{enumerate}

@@

\item Using probabilistic considerations only, show that the square of a stochastic matrix is also a
stochastic matrix.
\begin{quote}
If it was not a stochastic matrix, then there would be no way to represent the next state in the
system. The total probability of all the transitions must sum to 1, since there is a 100\% chance
of a transition occurring in every step of the chain.
\end{quote}

\item A stochastic matrix is doubly stochastic if the sum of the entries in each column is 1. If $A$
and $B$ are doubly stochastic square matrices, is $A^{2}$ doubly stochastic? Is $AB$?
\begin{quote}
Yes, the square of a doubly stochastic matrix is still a doubly stochastic matrix. $AB$ is also
still doubly stochastic.
\end{quote}

\item Write out inductive proofs for Theorems 2 and 3.

\item Use matrix multiplication to solve Exercise 8.
\[
\left(
\begin{array}{c c}
.5 & .5\\
.5 & .5
\end{array}
\right)
\cdot
\left(
\begin{array}{c c}
.5 & .5\\
.5 & .5
\end{array}
\right)
=
\]

\item Abigail bought a toy with her allowance this week (see Exercise 2). Find the probability that
she will buy a toy 4 weeks from now.

\item Find the distribution of birth weights (Example 1) after one generation if the initial
probability distribution is (.4, .3, .3).

\item Suppose the distribution of birth weights of a generation of daughters is
$p^{(1)}=7(0.31,0.45,0.24)$. Can you find the distribution of birth weights of the mothers?

\item Use the state diagram in Fig. 11.3 to find the probability that the process reaches state
$S_{3}$ in two steps if it starts in state $S_{1}$.

\item Let $P$ be the transition matrix of Eq. (4). Compute $P^{2}$. What can you say about the
first column of $P^{3}$ and $P^{n}$?

\setcounter{enumi}{20}
\item The matrices determined in Exercises 1-7 are all transition matrices for Markov processes.
Which ones are regular?
\begin{quote}
Exercises 1 and 2 are regular matrices, since they have all positive entries.
\end{quote}

\item Find, if possible, a fixed-point vector for each of the transition matrices of Exercises 1-7.
\begin{multicols}{2}
\begin{enumerate}
\item $\displaystyle \mathbf{w}=\left\{{5\over9}, {4\over9}\right\}$
\item $\displaystyle \mathbf{w}=\left\{{4\over7}, {3\over7}\right\}$
\item $\displaystyle \mathbf{w}=\left\{{23\over70}, {43\over70}, {2\over35}\right\}$
\item This is not a regular matrix, so no $\bf{w}$ exists.
\item This is not a regular matrix, so no $\bf{w}$ exists.
\item This is not a regular matrix, so no $\bf{w}$ exists.
\item This is not a regular matrix, so no $\bf{w}$ exists.
\end{enumerate}
\end{multicols}

\item Does a matrix of the form
$\displaystyle A=\left(\begin{array}{c c}a & b \\ -b & d\end{array}\right)$ have a fixed-point vector?
\begin{quote}
If it has a fixed-point vector, then that vector is given by
\[
\mathbf{w}=\left\{ {c\over b+c},{b\over b+c} \right\}
= \left\{ {-b\over b-b},{b\over b-b}\right\}
= \left\{ {-b\over 0},{b\over 0}\right\}.
\]
Therefore, there is no fixed-point vector for the matrix.
\end{quote}

\setcounter{enumi}{31}
\item Which of the transition matrices of Exercises 1-7 represent absorbing Markov processes?
\begin{quote}
Exercises 4 and 6 represent absorbing Markov processes, since they have nodes which are not
escapable.
\end{quote}
\end{enumerate}

### 2016-04-05 number_theory/hw/hw9.tex
working on number theory

\begin{enumerate}
\item Directly calculate $\displaystyle\sum\limits_{d|12}\phi(d)$ and verify that you obtain 12 as your
answer.
\begin{align*}
\sum_{d|12}\phi(d) &= \phi(1) + \phi(2) + \phi(3) + \phi(4) + \phi(6) + \phi(12)\\
&= 1 + 1 + 2 + 2 + 2 + 4\\
&= 12
\end{align*}
\end{enumerate}

\end{document}

>>>>

\begin{enumerate}
\item Directly calculate $\displaystyle\sum\limits_{d|12}\phi(d)$ and verify that you obtain 12 as your
answer.
\begin{align*}
\sum_{d|12}\phi(d) &= \phi(1) + \phi(2) + \phi(3) + \phi(4) + \phi(6) + \phi(12)\\
&= 1 + 1 + 2 + 2 + 2 + 4\\
&= 12
\end{align*}

\item Suppose that $p_{1},p_{2},\dots,p_{N}$ are distinct primes. Prove that
\[
{\phi(p_{1}p_{2}\dots p_{N}) \over p_{1}p_{2}\dots p_{N}}
=
\prod_{n=1}^{N}\left(1-{1\over p_{n}}\right)
\]
\begin{align*}
{\phi(p_{1}p_{2}\dots p_{N}) \over p_{1}p_{2}\dots p_{N}} &=
{\phi(p_{1})\phi(p_{2})\dots \phi(p_{N}) \over p_{1}p_{2}\dots p_{N}}\\
&= \prod_{n=1}^{N}{\phi(p_{n})\over p_{n}}\\
&= \prod_{n=1}^{N}{p_{n} -1 \over p_{n}}\\
&= \prod_{n=1}^{N}\left(1-{1\over p_{n}}\right)
\end{align*}

\item Find a value of $n$ such that $\phi(n)/n < 1/4$. What do you think is a good strategy for
choosing $n$ so that $\phi(n)/n$ is close to zero?
\begin{quote}
One value for which this holds true is $n=210$. $\phi(210)$ is 48, and $48/210 = 8/35 < 1/4$. One
strategy for finding these numbers would be to
\end{quote}

\item Suppose that $p$ is prime and $m$ and $n$ are non-negative integers.
\begin{enumerate}
\item Prove that $\phi(p^{m+n})\geq\phi(p^{m})\phi(p^{n})$.
\begin{quote}
We can consider $\phi(p^{m+n})$. If we let $m+n=j$, then we have $\phi(p^{j})$, which can be
expressed as $p^{j}-p^{j-1}$. If we consider $\phi(p^{m})\phi(p^{n})$, we have
\begin{align*}
\phi(p^{m})\phi(p^{n})&= (p^{m}-p^{m-1})(p^{n}-p^{n-1})\\
&= p^{m+n} - 2p^{m+n-1} + p^{m+n-2}\\
&= p^{j}-2p^{j-1}+p^{j-2}.
\end{align*}
If we compare the two values, we get
\begin{align*}
p^{j}-p^{j-1} &\stackrel{?}{\geq} p^{j}-2p^{j-1}+p^{j-2}\\
p^{j-1} &\geq p^{j-2}.
\end{align*}
We know this is true since $m,n>0$, so $j>0$.
\end{quote}

\item Under what additional assumptions on $m$ and $n$ do we obtain
$\phi(p^{m+n})=\phi(p^{m})\phi(p^{n})$?
\end{enumerate}

\end{enumerate}
\end{document}
